Towards a Verified Bzip2: A Functional Specification for Burrows--Wheeler

### 2019-05-06 draft_paper/draft_paper.tex
Finish appendices

Although it is difficult to prove, it is intuitive that prefix orders
that can look at more elements of the lists are ``stronger'' than
those that can look at fewer.

\begin{theorem}[key\_le\_firstn\_ge]
  For all $j$ and $k$ such that $k \ge j$,
  \begin{equation*}
    x \le_k y \implies x \le_j y
  \end{equation*}
\end{theorem}

@@

>>>>

We can describe the relationship between orders that can look at fewer
or more characters with the following two theorems. We will briefly
offer an explanation of why these theorems makes sense, but these
explanations do not mirror the Coq proofs we have for these theorems.

\begin{theorem}[key\_lt\_firstn\_ge]
  For all $j$ and $k$ such that $j \le k$,
  \begin{equation*}
    x <_j y \implies x <_k y
  \end{equation*}
\end{theorem}
\begin{proof}
  If $x <_j y$, then there must be some index $i < j$ such that $x_i <
  y_i$. Then $i < k$ as well, so $x <_k y$.
\end{proof}

\begin{theorem}[key\_le\_firstn\_ge]
  For all $j$ and $k$ such that $k \ge j$,
  \begin{equation*}
    x \le_k y \implies x \le_j y
  \end{equation*}
\end{theorem}
\begin{proof}
  If $x \le_k y$, then either the first $k$ characters are all
  equivalent, in which case the first $j$ characters are all
  equivalent as well, or there was some $i < k$ such that $x_i < y_i$,
  but at every index before $i$ $x$ and $y$ are equivalent. If $j <
  i$, then the first $j$ characters are still equivalent, and we have
  $x \equiv y \implies x \le_j y$. If $j \ge i$, then the deciding character is
  among the first $j$ and $x \le_j k$.
\end{proof}

@@

In this section, we prove the following theorem:
\begin{theorem}[hdsort\_sorted\_S]
  For all matrices $m$ and naturals $j$,
  \begin{equation*}
    \var{PrefixSorted}~j~(\var{map}~\var{tl}~m) \implies
    \var{PrefixSorted}~(j+1)~(\var{hdsort}~m)
  \end{equation*}
\end{theorem}

First we prove this lemma:
\begin{lemma}[insert\_Sorted\_S]
  For all matrices $m$, rows $r$, and naturals $j$,
  \begin{gather*}
    \var{PrefixSorted}~(j+1)~(\var{hdsort}~m) \implies \\
    \var{PrefixSorted}~j~(\var{tl}~r :: \var{map}~{tl}~m) \implies \\
    \var{PrefixSorted}~(j+1)~(\var{insert}_1~r~(\var{hdsort}~m))
  \end{gather*}
  where $\var{insert}_1$ is \var{insert} that uses the 1-prefix
  lexicographic ordering used by \var{hdsort}.
\end{lemma}
\begin{proof}
  We can destruct $\var{insert}_1~r~(\var{hdsort}~m)$ as $m_1 \dplus r
  :: m_2$, where $m_1 \dplus m_2 = \var{hdsort}~m$. Then by
  \var{Sorted\_app\_cons}, it suffices to show
  \begin{gather*}
    \var{PrefixSorted}~(j+1)~m_1 \land \\
    \var{PrefixSorted}~(j+1)~(r :: m_2) \land \\
    (\forall x \in m_1,~ r >_{j+1} x)
  \end{gather*}
  \begin{itemize}
  \item $\var{PrefixSorted}~(j+1)~m_1$ follows from the fact that
    $\var{hdsort}~m = m_1 \dplus m_2$ and
    $\var{PrefixSorted}~(j+1)~(\var{hdsort}~m)$.
  \item By similar reasoning, we know that
    $\var{PrefixSorted}~(j+1)~m_2$, so in order to show
    $\var{PrefixSorted}~(j+1)~(r :: m_2)$ all we need to prove is that
    for every row $x$ in $m_2$,
    \begin{equation*}
      r \le_{j+1} x
    \end{equation*}
    We know from the fact that $\var{insert}_1~r~(\var{hdsort}~m) =
    m_1 \dplus r :: m_2$ that $r \le_1 x$. Then by \var{key\_firstn\_S},
    we only need to show that
    \begin{equation*}
      \var{tl}~r \le_j \var{tl}~x
    \end{equation*}
    This follows from $\var{PrefixSorted}~j~(\var{tl}~r :: \var{map}~{tl}~m)$.
  \item We have that for any $x \in m_1$, $r >_1 x$ by the fact that
    $\var{insert}_1~r~(\var{hdsort}~m) = m_1 \dplus r :: m_2$. This
    gives us $r >_{j+1} x$ by \var{key\_lt\_firstn\_ge}. \qedhere
  \end{itemize}
\end{proof}

%% \begin{proof}
%%   We reason by induction on $m$. In the inductive case, we need to show
%%   \begin{equation*}
%%     \var{PrefixSorted}~(j+1)~(\var{hdsort}~(r::m))
%%   \end{equation*}
%%   If we simplify \var{hdsort}, this becomes
%%   \begin{equation*}
%%     \var{PrefixSorted}~(j+1)~(\var{insert}_1~r~(\var{hdsort}~m))
%%   \end{equation*}
%%   where $\var{insert}_1$ is \var{insert} using the $1$-prefix
%%   lexicographic ordering used in \var{hdsort}.
%%   We know from our assumption that
%%   \begin{equation*}
%%     \var{PrefixSorted}~j~(\var{tl}~x :: \var{map}~\var{tl}~m)
%%   \end{equation*}
%%   and from our induction hypothesis we have
%%   \begin{equation*}
%%     \var{PrefixSorted}~(j+1)~(\var{hdsort}~m)
%%   \end{equation*}
%% \end{proof}

  %% We can deconstruct $\var{insert}_1~r~(\var{hdsort}~m)$ as
  %% $m_1 \dplus r :: m_2$, where we also know that $m = m_1 \dplus m_2$.
  %% Now we have to show
  %% \begin{equation*}
  %%   \var{PrefixSorted}~(j+1)~(m_1 \dplus r :: m_2)
  %% \end{equation*}
